Comparing Power Optimizations in the Reorder Buffer

### 2011-06-08 paper/paper.tex
paper bg

\subsection{Latch ROB}
In addition to large power requirement to maintain the entries of the ROB, another 
large power consumer is the read ports of the ROB.  The read port of the ROB is used
to forward the results of an entry to the dispatch unit (DU) to use as an operand.
All though forwarding from the ROB occurs about 6\%\cite{kucuk} of all instructions 
in an N way super scalar processor there are 2N ports (to accommodate FP
 instructions) which add unnecessary complexity and power consumption for a small 
percentage of instructions.  It has been proposed by eliminating the read ports 
in the ROB all together, we can reduce energy consumption\cite{kucuk}.  However
without read ports, a small window appears where the data cannot be accessed,
starting from when the result is placed in to the ROB until the ROB commits that 
entry to the ARF.  To help mitigate stalls that might happen due to this hole, 
we introduce retention latches.

>>>>

\subsection{Latch ROB}
In addition to large power requirement to maintain the entries of the ROB, another 
large power consumer is the read ports of the ROB.  The read port of the ROB is used
to forward the results of an entry to the dispatch unit (DU) to use as an operand.
All though forwarding from the ROB occurs about 6\%\cite{kucuk} of all instructions 
in an N way super scalar processor there are 2N ports (to accommodate FP
 instructions) which add unnecessary complexity and power consumption for a small 
percentage of instructions.  It has been proposed by eliminating the read ports 
in the ROB all together, we can reduce energy consumption\cite{kucuk}.  However
without read ports, a small window appears where the data cannot be accessed,
starting from when the result is placed in to the ROB until the ROB commits that 
entry to the ARF.  To help mitigate stalls that might happen due to this hole, 
we introduce retention latches.   Retention latches are small memory elements and 
we use them to cache the most recent results that have been written into the ROB.
This way instructions are still able to access the most recently produced results
and not stall.  However if the operand needed is not in the ROB, the stream will 
stall until a commit is made to the ARF, and that result value is forwarded to 
the dispatch unit immediately.
